The paper on \Large \bf Blockchain Frontrunning

### 2018-09-12 frontrunning/Latex/sections/paper.tex
[frontrunning] fix subsections + one fast round of typo fixing

%The two significant factors which cause the front-running practice to happen within the financial markets are (i) imperfect competition and (ii) liquidity uncertainty~\cite{liang2005distressed}. Any sort of front-running activity wis considered as an unethical and illegal practice as it is unfairly beneficial for a few system participants who have the privilege of acting on this information and taking advantages at the expense of the investors.

\subsection{How are Blockchain Applications Vulnerable to Front-running?} %which blockchain applications?
\label{sec:Front Running on the Blockchains}

%Jeremy's Comments: Find some examples; you have to justify why you talked about only a specific group of examples. Need to have a criteria for what is vulnerable to frontrunning: like top ten dapps from specific blogs. Have a criteria.

@@

Malinova and Park discuss different design settings for financial markets and trading platforms on decentral ledgers ~\cite{malinova2017market}.  They argue that blockchain brings a new options of transparency which affects traders behaviour, specially in the presence of intermediary front-runners. They model indirect trading costs and use the economic literature on search and trading at decentralized exchanges in their design settings to make front-running costly and inefficient.

@@

One of the properties of Blockchain technology was to overcome double-spend, however this does not apply for fast payments and zero-confirmation transactions. \textit{Double-spend attack}, as the name implies, the adversary spends a coin more than one time. To do so, he would use the coin as an input of a transaction and broadcast it to the network, before the transaction gets validated and executed, he broadcasts another transaction associated with a higher transaction fee which spends the same input ~\cite{bamert2013have, karame2012double}. This adversarial behaviour is similar to the front-running attack where the attacker would front-run other users by paying a higher transaction fee when broadcasting his transactions to the network, hence a higher priority transaction for miners to include in the block they mine.

@@

\section{Status ICO} 
As discussed in~\ref{sec:Front Running on the Blockchains}, capped ICOs are considerable applications to be investigated in terms of the front-running attacks, Status ICO is such application. In June 2017, Status.im~\cite{statuswhitepaper} started its ICO and reached its cap within 3 hours, resulted in close to 300,000 Ether of funding. In order to prevent large investments to take place early and limit the amount of Ether deposited in an investment, the team at the Status ICO used a \textit{fair} token distribution method called \textit{dynamic ceiling} as an attempt to increase the time window for smaller investors. They implemented multiple caps --ceiling-- which each had a maximum amount that could be deposited. In this case, every deposit was checked by the smart contract and the exceeding amount was refunded to the sender while the accepted amount was sent to the Status.im wallet ~\cite{statusicoanalysis}.

During the time the ICO was open for participation, there were reports of Ethereum network being unusable and transactions were not confirming. Further study showed that some mining pools might have been manipulating the network for their own profit. In addition, there were many transactions sent with higher gas price to front-run other transactions, however these transactions were failed due to the restriction in the ICO smart contract to reject any transactions with higher than 50GWei gas price (as an attempt to prevent the aforementioned front-running).

@@

Ethereum blockchain contains all transaction ever made on Ethereum, however, it is not easy to see the behaviour of the participants and miners solely using the default clients nor online blockchain explorers. In order to get the required data, Ethereum blockchain data should be converted to a queryable format. For this case study, we used open source projects such as Go Ethereum implementation\footnote{Official Go implementation of the Ethereum protocol\url{https://github.com/ethereum/go-ethereum}}, python script for extracting, transforming and loading ethereum blocks, named \texttt{ethereum-etl}~\cite{ethereumetl} and Google BigQuery\footnote{A fast, highly scalable, cost-effective, and fully managed cloud data warehouse for analytics \url{https://cloud.google.com/bigquery/}}. Using this software stack, we were able to query for all the transactions and blocks within the window of Status.im ICO for further analysis which was mainly done by data analysis tool \texttt{Tableau}\footnote{Tableau is a data visualization software that is used for data science and business intelligence.\url{https://www.tableau.com/}}. A copy of this dataset and the findings can be found in \texttt{github.com / ANONYMOUS}.

@@

\subsubsection{Exchanges}
Exchanges such as EtherDelta\footnote{also known as ForkDelta for the UI \url{https://forkdelta.app/}}, are closest form to decentral exchanges, however their orderbooks are stored on a central server they control and shown on their user interface (website). Other than the fact that they can re-order or censor the orders on their servers, there are other risks of front-running nature in these designs. Similarly 0xProtocol uses \textit{Relayers} which act as the orderbook holders and are not prune to cheating. One main issue with these exchanges is \textit{cancellation grief}. Although the orderbooks are not in the smart contract, in order to prevent denial of service attacks, user needs to send a transaction to cancel her order. In this case, when an adversarial actor sees the cancellation transaction, he sends a fill order transactions with higher gasPrice to get in front of the cancellation order and take the order before it is cancelled. As mentioned in ~\ref{minersfrontrunning}, miners are in a better control to perform such attack.

@@

\subsubsection{Domain Name Front-running} \hfill\\

\noindent As mentioned in Section~\ref{sec:Front Running on the Blockchains}, although front-running attacks have been more showcased in the context of decentralized exchanges and trading systems, they are not yet limited to the financial markets. Front-running can also occur within other (non-financial) blockchain applications such as naming systems. Recently, many blockchain-based domain registrars have been designed and implemented to eliminate the role of central parties \ie domain name system (DNS) which itself introduces single point of failures in the entire web. Ethereum Name Service (ENS)\footnote{\url{https://ens.domains/}}, is the most popular decentral naming service that uses a sealed auction to facilitate the procedures of purchasing new \texttt{.eth} domain names. These domains which can be resolved using any Ethereum web3 compatible browser. Although a sealed auction (\textit{startAuctionsAndBid}) is used while implementing the ENS, the deposit should be equal or bigger than the bid, this information leakage could be used to front-run the transaction with the same bid amount. By doing so the front-runner invalidates the original bid and hence registers the domain name himself. Ghazal is another Ethereum-based naming system that allows users to register \texttt{.ghazal} domain names and bind certificates to those names~\cite{moosavighazal}.  In Ghazal model, a user would register domain name by executing the \textit{registerdomain} function from the Ghazal smart contract with the domain name in plain text as the function argument. As mentioned before, These transactions will sit in the mempool so that it would be mined by Ethereum miners and included in the block. During this period in which the transaction is not yet confirmed, front-running attack can occur by (i) dishonest miners and/or (ii) regular non-miner user. In the first case, a miner would intercept the \textit{registerdomain} transaction and register that name ahead of the user. A regular non-miner node in the Ethereum network can front-run other user's \textit{registerdomain} at a good profit by paying higher transaction fees. This practice allows the adversarial party  to register an unregistered domain immediately after someone searched for it, and sell the domain name to the users for a higher price. Note that, domain name front-running also occurs in the vanilla system. In this case, front-runners would monitor and identify domain names of interest, and sell it back to the other users with a higher price~\cite{sac022en33:online}. Front-running practice can happen in the traditional domain name system (DNS) as well. Domain name registrars can steal available domain names once seeing a query from users that looks up that domain's availability. In this case, the registrar can register the domain name ahead and sell it back to the user for a higher price.

@@

%______________________________________________________________%
\subsection{Crypto-Collectibles Games} \hfill\\

Crypto-Collectibles, also known as ERC-721 Non-Fungible Token Standard~\cite{erc721}, are one of the applications of Ethereum blockchain. Cryptokitties is known to be one of the first and the most popular use cases of ERC721 tokens~\cite{cryptokitties}. Their marketcap reached more than 6 million dollars in the first months their launch. Crypto-Collectibles, are similar to ERC-20 tokens and can be listed on exchanges. Most of the front-running issues applied to the exchanges also apply to DApps using ERC-721 tokens. \\

@@

\subsection{Gambling} \hfill\\ 
Due to deterministic nature of blockchains, running gambling games based on randomness is close to impossible. There had been attempts to use block headers or such as the seed to randomness but almost all have failed as every data accessible to the smart contract at the time of execution, is also accessible to other smart contracts and they could predict the output on the run, thus prone to front-running and miner's manipulation. Many infamous gambling games are now running on Ethereum network and front-running them requires a level of sophistication and knowledge of the internal of the games and Ethereum.\\
One of the most popular smart contracts on Ethereum blockchain belongs to a controversial gambling game named \texttt{Fomo3d}, also known as \textit{Exit Scam}\footnote{\url{https://exitscam.me/play}}. The aim of this game is to be the last person purchasing the key when the time goes to zero in which every ticket purchase increases the timer by 30 seconds. It was believed that because of the nature of the game, it will never end, however on August 22, 2018 the first round of the game ended and the winner took 10,469 Ether\footnote{First winner of Fomo3D, won 10,469 Ether \url{ https://etherscan.io/tx/0xe08a519c03cb0aed0e04b33104112d65fa1d3a48cd3aeab65f047b2abce9d508}}, equivalent to 2.1 Million dollars on the time of writing. 
It is believed that the winner, used a sophisticated approach to finish last~\cite{fomo3dhacker}. The winner deployed a few high gas consumption smart contracts prior to the win, when the timer of the game reached about 3 minutes, the winner bought 1 ticket and sent multiple high gasPrice transactions to aforementioned smart contracts to lure the miners to mine these highly profitable transactions. Luckily enough the winner was able to congest the network in a way that no other ticket purchases were mined in the 3 minutes window and the miners were focused on the high gasPriced transactions. All the details mentioned here are drawn from the transaction history on the blockchain and unless otherwise proven seem to be the way the winner acted to win the pot. 
%probably needs more work to be good enough. more resources will be out in the coming days !

@@

In the traditional markets, front-running is considered unethical and also illegal. In the blockchain space, the designers of the decentral applications cannot rely on the justice system for unethical behaviour, and they must assume that each participant in the network acts rationally in their own self interest, the application will be operating in an adversarial environment~\cite{0xfrontrunning:online}.
A few decentral exchange projects such as \textit{EtherDelta} and \textit{0xProject}~\cite{warren20170x}, have proposed solutions for front-running, which is off-chain order books. The methods discussed within these projects prevent blockchain network participants from front running the orders as the orders are private and not known to the network and will not be broadcasted. However they are introducing third parties, \eg relayers in 0xproject, to be managing the orders with the promise not to reorder or front-run other orders.

@@

The details of how Submarine commitments work is outside the scope of this paper, however in summary, using the way Ethereum transactions are constructed and ECDSA ECRecover functionality it is possible to generate one outgoing transaction from an address in which no private key exists. By constructing a transaction from that address to the DApp and funding that address we fulfil the commitment phase and by revealing the details and broadcasting the constructed transaction we reveal the commitment ~\cite{libsubmarine}, see Figure~\ref{fig:LibSubmarine}.  Drawbacks of this scheme are similar to ~\ref{CommitmentScheme}, with an addition of two more transactions to finalize the functionality, which results in more cost and complicated user experience.

>>>>

\subsection{How are Blockchain Applications Vulnerable to Front-running?} 
\label{sec:Front Running on the Blockchains}

@@

Malinova and Park discuss different design settings for financial markets and trading platforms on decentral ledgers ~\cite{malinova2017market}.  They argue that blockchain brings a new options of transparency which affects traders behavior, specially in the presence of intermediary front-runners. They model indirect trading costs and use the economic literature on search and trading at decentralized exchanges in their design settings to make front-running costly and inefficient.

@@

One of the properties of Blockchain technology was to overcome double-spend, however this does not apply for fast payments and zero-confirmation transactions. \textit{Double-spend attack}, as the name implies, the adversary spends a coin more than one time. To do so, he would use the coin as an input of a transaction and broadcast it to the network, before the transaction gets validated and executed, he broadcasts another transaction associated with a higher transaction fee which spends the same input ~\cite{bamert2013have, karame2012double}. This adversarial behavior is similar to the front-running attack where the attacker would front-run other users by paying a higher transaction fee when broadcasting his transactions to the network, hence a higher priority transaction for miners to include in the block they mine.

@@

\section{Status ICO} 
As discussed in~\ref{sec:Front Running on the Blockchains}, capped ICOs are considerable applications to be investigated in terms of the front-running attacks, Status ICO is such application. In June 2017, \textit{Status.im}~\cite{statuswhitepaper} started its ICO and reached its cap within 3 hours, resulted in close to 300,000 Ether of funding. In order to prevent large investments to take place early and limit the amount of Ether deposited in an investment, the team at the Status ICO used a \textit{fair} token distribution method called \textit{dynamic ceiling} as an attempt to increase the time window for smaller investors. They implemented multiple caps --ceiling-- which each had a maximum amount that could be deposited. In this case, every deposit was checked by the smart contract and the exceeding amount was refunded to the sender while the accepted amount was sent to the Status.im wallet ~\cite{statusicoanalysis}.

During the time the ICO was open for participation, there were reports of Ethereum network being unusable and transactions were not confirming. Further study showed that some mining pools might have been manipulating the network for their own profit. In addition, there were many transactions sent with higher gas price to front-run other transactions, however these transactions were failed due to the restriction in the ICO smart contract to reject any transactions with higher than 50 \textit{GWei} gas price (as an attempt to prevent the aforementioned front-running).

@@

Ethereum blockchain contains all transaction ever made on Ethereum, however, it is not easy to see the behavior of the participants and miners solely using the default clients nor online blockchain explorers. In order to get the required data, Ethereum blockchain data should be converted to a queryable format. For this case study, we used open source projects such as Go Ethereum implementation\footnote{Official Go implementation of the Ethereum protocol\url{https://github.com/ethereum/go-ethereum}}, python script for extracting, transforming and loading ethereum blocks, named \texttt{ethereum-etl}~\cite{ethereumetl} and Google BigQuery\footnote{A fast, highly scalable, cost-effective, and fully managed cloud data warehouse for analytics \url{https://cloud.google.com/bigquery/}}. Using this software stack, we were able to query for all the transactions and blocks within the window of Status.im ICO for further analysis which was mainly done by data analysis tool \texttt{Tableau}\footnote{Tableau is a data visualization software that is used for data science and business intelligence.\url{https://www.tableau.com/}}. A copy of this dataset and the findings can be found in \texttt{github.com / ANONYMOUS}.

@@

\subsubsection{Exchanges.}
Exchanges such as EtherDelta\footnote{also known as ForkDelta for the UI \url{https://forkdelta.app/}}, are closest form to decentral exchanges, however their orderbooks are stored on a central server they control and shown on their user interface (website). Other than the fact that they can re-order or censor the orders on their servers, there are other risks of front-running nature in these designs. Similarly 0xProtocol uses \textit{Relayers} which act as the orderbook holders and are not prune to cheating. One main issue with these exchanges is \textit{cancellation grief}. Although the orderbooks are not in the smart contract, in order to prevent denial of service attacks, user needs to send a transaction to cancel her order. In this case, when an adversarial actor sees the cancellation transaction, he sends a fill order transactions with higher gasPrice to get in front of the cancellation order and take the order before it is cancelled. As mentioned in ~\ref{minersfrontrunning}, miners are in a better control to perform such attack.

@@

\subsubsection{Name Services.}
\noindent As mentioned in Section~\ref{sec:Front Running on the Blockchains}, although front-running attacks have been more showcased in the context of decentralized exchanges and trading systems, they are not yet limited to the financial markets. Front-running can also occur within other (non-financial) blockchain applications such as naming systems. Recently, many blockchain-based domain registrars have been designed and implemented to eliminate the role of central parties \ie domain name system (DNS) which itself introduces single point of failures in the entire web. Ethereum Name Service (ENS)\footnote{\url{https://ens.domains/}}, is the most popular decentral naming service that uses a sealed auction to facilitate the procedures of purchasing new \texttt{.eth} domain names. These domains which can be resolved using any Ethereum web3 compatible browser. Although a sealed auction (\textit{startAuctionsAndBid}) is used while implementing the ENS, the deposit should be equal or bigger than the bid, this information leakage could be used to front-run the transaction with the same bid amount. By doing so the front-runner invalidates the original bid and hence registers the domain name himself. Ghazal is another Ethereum-based naming system that allows users to register \texttt{.ghazal} domain names and bind certificates to those names~\cite{moosavighazal}.  In Ghazal model, a user would register domain name by executing the \textit{registerdomain} function from the Ghazal smart contract with the domain name in plain text as the function argument. As mentioned before, These transactions will sit in the mempool so that it would be mined by Ethereum miners and included in the block. During this period in which the transaction is not yet confirmed, front-running attack can occur by (i) dishonest miners and/or (ii) regular non-miner user. In the first case, a miner would intercept the \textit{registerdomain} transaction and register that name ahead of the user. A regular non-miner node in the Ethereum network can front-run other user's \textit{registerdomain} at a good profit by paying higher transaction fees. This practice allows the adversarial party  to register an unregistered domain immediately after someone searched for it, and sell the domain name to the users for a higher price. Note that, domain name front-running also occurs in the vanilla system. In this case, front-runners would monitor and identify domain names of interest, and sell it back to the other users with a higher price~\cite{sac022en33:online}. Front-running practice can happen in the traditional domain name system (DNS) as well. Domain name registrars can steal available domain names once seeing a query from users that looks up that domain's availability. In this case, the registrar can register the domain name ahead and sell it back to the user for a higher price.

@@

%______________________________________________________________%
\subsubsection{Crypto-Collectibles Games.}

Crypto-Collectibles, also known as ERC-721 Non-Fungible Token Standard~\cite{erc721}, are one of the applications of Ethereum blockchain. Cryptokitties is known to be one of the first and the most popular use cases of ERC721 tokens~\cite{cryptokitties}. Their market cap reached more than 6 million dollars in the first months their launch. Crypto-Collectibles, are similar to ERC-20 tokens and can be listed on exchanges. Most of the front-running issues applied to the exchanges also apply to DApps using ERC-721 tokens. \\

@@

\subsubsection{Gambling.} 
Due to deterministic nature of blockchains, running gambling games based on randomness is close to impossible. There had been attempts to use block headers or such as the seed to randomness but almost all have failed as every data accessible to the smart contract at the time of execution, is also accessible to other smart contracts and they could predict the output on the run, thus prone to front-running and miner's manipulation. Many infamous gambling games are now running on Ethereum network and front-running them requires a level of sophistication and knowledge of the internal of the games and Ethereum.\\
One of the most popular smart contracts on Ethereum blockchain belongs to a controversial gambling game named \texttt{Fomo3d}, also known as \textit{Exit Scam}\footnote{\url{https://exitscam.me/play}}. The aim of this game is to be the last person purchasing the key when the time goes to zero in which every ticket purchase increases the timer by 30 seconds. It was believed that because of the nature of the game, it will never end, however on August 22, 2018 the first round of the game ended and the winner took 10,469 Ether\footnote{First winner of Fomo3D, won 10,469 Ether \url{ https://etherscan.io/tx/0xe08a519c03cb0aed0e04b33104112d65fa1d3a48cd3aeab65f047b2abce9d508}}, equivalent to 2.1 Million dollars on the time of writing. 
It is believed that the winner, used a sophisticated approach to finish last~\cite{fomo3dhacker}. The winner deployed a few high gas consumption smart contracts prior to the win, when the timer of the game reached about 3 minutes, the winner bought 1 ticket and sent multiple high gasPrice transactions to aforementioned smart contracts to lure the miners to mine these highly profitable transactions. Luckily enough the winner was able to congest the network in a way that no other ticket purchases were mined in the 3 minutes window and the miners were focused on the high gasPriced transactions. All the details mentioned here are drawn from the transaction history on the blockchain and unless otherwise proven seem to be the way the winner acted to win the pot. 
%probably needs more work to be good enough. more resources will be out in the coming days !

@@

In the traditional markets, front-running is considered unethical and also illegal. In the blockchain space, the designers of the decentral applications cannot rely on the justice system for unethical behavior, and they must assume that each participant in the network acts rationally in their own self interest, the application will be operating in an adversarial environment~\cite{0xfrontrunning:online}.
A few decentral exchange projects such as \textit{EtherDelta} and \textit{0xProject}~\cite{warren20170x}, have proposed solutions for front-running, which is off-chain order books. The methods discussed within these projects prevent blockchain network participants from front running the orders as the orders are private and not known to the network and will not be broadcasted. However they are introducing third parties, \eg relayers in 0xproject, to be managing the orders with the promise not to reorder or front-run other orders.

@@

The details of how Submarine commitments work is outside the scope of this paper, however in summary, using the way Ethereum transactions are constructed and ECDSA ECRecover functionality it is possible to generate one outgoing transaction from an address in which no private key exists. By constructing a transaction from that address to the DApp and funding that address we fulfill the commitment phase and by revealing the details and broadcasting the constructed transaction we reveal the commitment ~\cite{libsubmarine}, see Figure~\ref{fig:LibSubmarine}.  Drawbacks of this scheme are similar to ~\ref{CommitmentScheme}, with an addition of two more transactions to finalize the functionality, which results in more cost and complicated user experience.
